Data Exploration

### 2019-02-04 src/data_exploration.Rmd
monthly spearmans rerun with selected data + minor alteration to the folders output sent to

looks like not much difference... what about a kruskal wallace test?

@@

```{r}
vif_allpresab <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allpresab, "../output/bio/vif_allpresab.csv", row.names = TRUE)
vif_allpresab
```
interpret - see https://stats.stackexchange.com/questions/70679/which-variance-inflation-factor-should-i-be-using-textgvif-or-textgvif/96584#96584
To make GVIFs comparable across dimensions, we suggested using GVIF^(1/(2*Df)), where Df is the number of coefficients in the subset (ref fox and motette 1992 in zotero)
or the 2 continuous variables, GVIFˆ(1/(2*Df)) (which is basically the square root of the VIF/GVIF value as DF=1) is the proportional change of the standard error and confidence interval of their coefficients due to the level of collinearity. The GVIFˆ(1/(2*Df)) value of the categorical variable is a similar measure for the reduction in precision of the coefficients' estimation due to collinearity. 
apparently i just need to square GVIF^(1/(2*Df)) and then use the normal VIF "rule of thumb"...

```{r}
vif_allpresab_sq <- read.csv("../output/bio/vif_allpresab.csv", header = TRUE)
vif_allpresab_sq$GVIF2Dfsq <- vif_allpresab_sq$GVIF..1..2.Df..^2
write.csv(vif_allpresab_sq, "../output/bio/vif_allpresab_sq.csv", row.names = TRUE)
vif_allpresab_sq
```

@@

```{r vif all but gear}
vif_allbutgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutgear, "../output/bio/vif_allbutgear.csv", row.names = TRUE)
vif_allbutgear
```

As per SLR suggestion, rerun without gear and most highly correlated variables

```{r vif without gear + high corr}
vif_allbutgearhighlycorr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbutgearhighlycorr, "../output/bio/vif_allbutgearhighlycorr.csv", row.names = TRUE)
vif_allbutgearhighlycorr
```

@@

```{r vif all but high corr}
vif_allbuthighlycorr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter + gear, data = presab))
write.csv(vif_allbuthighlycorr, "../output/bio/vif_allbuthighlycorr.csv", row.names = TRUE)
vif_allbuthighlycorr

vif_allbuthighlycorr_sq <- read.csv("../output/bio/vif_allbuthighlycorr.csv", header = TRUE)
vif_allbuthighlycorr_sq$GVIF2Dfsq <- vif_allbuthighlycorr_sq$GVIF..1..2.Df..^2
write.csv(vif_allbuthighlycorr_sq, "../output/bio/vif_allbuthighlycorr_sq.csv", row.names = TRUE)
vif_allbuthighlycorr_sq
```

@@

remove amo_prev
```{r vif all but amoprev}
vif_allbutamo_prev <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbutamo_prev, "../output/bio/vif_allbutamo_prev.csv", row.names = TRUE)
vif_allbutamo_prev

vif_allbutamo_prev_sq <- read.csv("../output/bio/vif_allbutamo_prev.csv", header = TRUE)
vif_allbutamo_prev_sq$GVIF2Dfsq <- vif_allbutamo_prev_sq$GVIF..1..2.Df..^2
write.csv(vif_allbutamo_prev_sq, "../output/bio/vif_allbutamo_prev_sq.csv", row.names = TRUE)
vif_allbutamo_prev_sq
```
and leave gear out

remove amo_prev + gear
```{r vif all but amoprev + gear}
vif_allbutamo_prevgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbutamo_prevgear, "../output/bio/vif_allbutamo_prevgear.csv", row.names = TRUE)
vif_allbutamo_prevgear
```

remove chl_depth
```{r vif all but chldepth}
vif_allbutchl_depth <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutchl_depth, "../output/bio/vif_allbutchl_depth.csv", row.names = TRUE)
vif_allbutchl_depth

vif_allbutchl_depth_sq <- read.csv("../output/bio/vif_allbutchl_depth.csv", header = TRUE)
vif_allbutchl_depth_sq$GVIF2Dfsq <- vif_allbutchl_depth_sq$GVIF..1..2.Df..^2
write.csv(vif_allbutchl_depth_sq, "../output/bio/vif_allbutchl_depth_sq.csv", row.names = TRUE)
vif_allbutchl_depth_sq
```

remove chl_depth and gear
```{r vif all but chldepth+gear}
vif_allbutchl_depthgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutchl_depthgear, "../output/bio/vif_allbutchl_depthgear.csv", row.names = TRUE)
vif_allbutchl_depthgear
```

remove ssh_surface
```{r vif all but ssh}
vif_allbutssh_surface <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutssh_surface, "../output/bio/vif_allbutssh_surface.csv", row.names = TRUE)
vif_allbutssh_surface

vif_allbutssh_surface_sq <- read.csv("../output/bio/vif_allbutssh_surface.csv", header = TRUE)
vif_allbutssh_surface_sq$GVIF2Dfsq <- vif_allbutssh_surface_sq$GVIF..1..2.Df..^2
write.csv(vif_allbutssh_surface_sq, "../output/bio/vif_allbutssh_surface_sq.csv", row.names = TRUE)
vif_allbutssh_surface_sq
```

remove ssh_surface & gear
```{r vif all but ssh+gear}
vif_allbutssh_surfacegear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutssh_surfacegear, "../output/bio/vif_allbutssh_surfacegear.csv", row.names = TRUE)
vif_allbutssh_surfacegear
```

remove o2_surface
```{r vif all but o2surface}
vif_allbuto2_surface <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbuto2_surface, "../output/bio/vif_allbuto2_surface.csv", row.names = TRUE)
vif_allbuto2_surface

vif_allbuto2_surface_sq <- read.csv("../output/bio/vif_allbuto2_surface.csv", header = TRUE)
vif_allbuto2_surface_sq$GVIF2Dfsq <- vif_allbuto2_surface_sq$GVIF..1..2.Df..^2
write.csv(vif_allbuto2_surface_sq, "../output/bio/vif_allbuto2_surface_sq.csv", row.names = TRUE)
vif_allbuto2_surface_sq
```

remove o2_surface & gear
```{r vif all but o2surface+gear}
vif_allbuto2_surfacegear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbuto2_surfacegear, "../output/bio/vif_allbuto2_surfacegear.csv", row.names = TRUE)
vif_allbuto2_surfacegear
```

@@

remove salinity_surface only
```{r vif all but salsurface}
vif_allbutsalinitysurface <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutsalinitysurface, "../output/bio/vif_allbutsalinitysurface.csv", row.names = TRUE)
vif_allbutsalinitysurface

vif_allbutsalinitysurface_sq <- read.csv("../output/bio/vif_allbutsalinitysurface.csv", header = TRUE)
vif_allbutsalinitysurface_sq$GVIF2Dfsq <- vif_allbutsalinitysurface_sq$GVIF..1..2.Df..^2
write.csv(vif_allbutsalinitysurface_sq, "../output/bio/vif_allbutsalinitysurface_sq.csv", row.names = TRUE)
vif_allbutsalinitysurface_sq
```
and without gear

```{r vif all but salsurface+gear}
vif_allbutsalinitysurfacegear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutsalinitysurfacegear, "../output/bio/vif_allbutsalinitysurfacegear.csv", row.names = TRUE)
vif_allbutsalinitysurfacegear
```

now remove temp_surface plus highly correlated
```{r vif allbut high corr + temp surface/gear/nogear}
#with gear
vif_allbuthighcorrtempsurface <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbuthighcorrtempsurface, "../output/bio/vif_allbuthighcorrtempsurface.csv", row.names = TRUE)
vif_allbuthighcorrtempsurface

vif_allbuthighcorrtempsurface_sq <- read.csv("../output/bio/vif_allbuthighcorrtempsurface.csv", header = TRUE)
vif_allbuthighcorrtempsurface_sq$GVIF2Dfsq <- vif_allbuthighcorrtempsurface_sq$GVIF..1..2.Df..^2
write.csv(vif_allbuthighcorrtempsurface_sq, "../output/bio/vif_allbuthighcorrtempsurface_sq.csv", row.names = TRUE)
vif_allbuthighcorrtempsurface_sq

#without gear
vif_allbuthighcorrtempsurfacegear <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbuthighcorrtempsurfacegear, "../output/bio/vif_allbuthighcorrtempsurfacegear.csv", row.names = TRUE)
vif_allbuthighcorrtempsurfacegear
```

now remove temp_surface plus salinity_surface
```{r vif allbut temp + salinity surface/gear/nogear}
#with gear
vif_allbuttempsalsurface <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbuttempsalsurface, "../output/bio/vif_allbuttempsalsurface.csv", row.names = TRUE)
vif_allbuttempsalsurface

vif_allbuttempsalsurface_sq <- read.csv("../output/bio/vif_allbuttempsalsurface.csv", header = TRUE)
vif_allbuttempsalsurface_sq$GVIF2Dfsq <- vif_allbuttempsalsurface_sq$GVIF..1..2.Df..^2
write.csv(vif_allbuttempsalsurface_sq, "../output/bio/vif_allbuttempsalsurface_sq.csv", row.names = TRUE)
vif_allbuttempsalsurface_sq

#without gear
vif_allbuttempsalsurfacegear <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbuttempsalsurfacegear, "../output/bio/vif_allbuttempsalsurfacegear.csv", row.names = TRUE)
vif_allbuttempsalsurfacegear
```

now remove temp_surface plus salinity_surface + highly correlated
```{r vif allbut high corr +  temp + salinity surface/gear/nogear}
#with gear
vif_allbuthighcorrtempsalsurface <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbuthighcorrtempsalsurface, "../output/bio/vif_allbuthighcorrtempsalsurface.csv", row.names = TRUE)
vif_allbuthighcorrtempsalsurface

vif_allbuthighcorrtempsalsurface_sq <- read.csv("../output/bio/vif_allbuthighcorrtempsalsurface.csv", header = TRUE)
vif_allbuthighcorrtempsalsurface_sq$GVIF2Dfsq <- vif_allbuthighcorrtempsalsurface_sq$GVIF..1..2.Df..^2
write.csv(vif_allbuthighcorrtempsalsurface_sq, "../output/bio/vif_allbuthighcorrtempsalsurface_sq.csv", row.names = TRUE)
vif_allbuthighcorrtempsalsurface_sq

#without gear
vif_allbuthighcorrtempsalsurfacegear <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbuthighcorrtempsalsurfacegear, "../output/bio/vif_allbuthighcorrtempsalsurfacegear.csv", row.names = TRUE)
vif_allbuthighcorrtempsalsurfacegear
```

@@

```{r}
vif_allbutnaoamo <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear, data = presab))
write.csv(vif_allbutnaoamo, "../output/bio/vif_allbutnaoamo.csv", row.names = TRUE)
vif_allbutnaoamo

>>>>

what about a kruskal wallace test?

@@

```{r}
vif_allpresab <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allpresab, "../output/bio/vif/vif_allpresab.csv", row.names = TRUE)
vif_allpresab
```
interpret - see https://stats.stackexchange.com/questions/70679/which-variance-inflation-factor-should-i-be-using-textgvif-or-textgvif/96584#96584
To make GVIFs comparable across dimensions, we suggested using GVIF^(1/(2*Df)), where Df is the number of coefficients in the subset (ref fox and motette 1992 in zotero)
or the 2 continuous variables, GVIFˆ(1/(2*Df)) (which is basically the square root of the VIF/GVIF value as DF=1) is the proportional change of the standard error and confidence interval of their coefficients due to the level of collinearity. The GVIFˆ(1/(2*Df)) value of the categorical variable is a similar measure for the reduction in precision of the coefficients' estimation due to collinearity. 
apparently i just need to square GVIF^(1/(2*Df)) and then use the normal VIF "rule of thumb"...

```{r}
vif_allpresab_sq <- read.csv("../output/bio/vif/vif_allpresab.csv", header = TRUE)
vif_allpresab_sq$GVIF2Dfsq <- vif_allpresab_sq$GVIF..1..2.Df..^2
write.csv(vif_allpresab_sq, "../output/bio/vif/vif_allpresab.csv", row.names = TRUE)
vif_allpresab_sq
```

@@

```{r vif all but gear}
vif_allbutgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutgear, "../output/bio/vif/vif_allbutgear.csv", row.names = TRUE)
vif_allbutgear
```
xx
As per SLR suggestion, rerun without gear and most highly correlated variables

```{r vif without gear + high corr}
vif_allbutgearhighlycorr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbutgearhighlycorr, "../output/bio/vif/vif_allbutgearhighlycorr.csv", row.names = TRUE)
vif_allbutgearhighlycorr
```

@@

```{r vif all but high corr}
vif_allbuthighlycorr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter + gear, data = presab))
write.csv(vif_allbuthighlycorr, "../output/bio/vif/vif_allbuthighlycorr.csv", row.names = TRUE)
vif_allbuthighlycorr

vif_allbuthighlycorr_sq <- read.csv("../output/bio/vif/vif_allbuthighlycorr.csv", header = TRUE)
vif_allbuthighlycorr_sq$GVIF2Dfsq <- vif_allbuthighlycorr_sq$GVIF..1..2.Df..^2
write.csv(vif_allbuthighlycorr_sq, "../output/bio/vif/vif_allbuthighlycorr.csv", row.names = TRUE)
vif_allbuthighlycorr_sq
```

@@

remove bottom_depth
```{r vif all but botdepth}
vif_allbutbot_prev <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutbot_prev, "../output/bio/vif/vif_allbutbot_prev.csv", row.names = TRUE)

vif_allbutbot_prev <- read.csv("../output/bio/vif/vif_allbutbot_prev.csv", header = TRUE)
vif_allbutbot_prev$GVIF2Dfsq <- vif_allbutbot_prev$GVIF..1..2.Df..^2
write.csv(vif_allbutbot_prev, "../output/bio/vif/vif_allbutbot_prev.csv", row.names = TRUE)
vif_allbutbot_prev
```

and leave gear out

remove bottom_depth + gear
```{r vif all but botddepth + gear}
vif_allbutbot_prevgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutbot_prevgear, "../output/bio/vif/vif_allbutbot_prevgear.csv", row.names = TRUE)
vif_allbutbot_prevgear
```

remove amo_prev
```{r vif all but amoprev}
vif_allbutamo_prev <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbutamo_prev, "../output/bio/vif/vif_allbutamo_prev.csv", row.names = TRUE)

vif_allbutamo_prev <- read.csv("../output/bio/vif/vif_allbutamo_prev.csv", header = TRUE)
vif_allbutamo_prev$GVIF2Dfsq <- vif_allbutamo_prev$GVIF..1..2.Df..^2
write.csv(vif_allbutamo_prev, "../output/bio/vif/vif_allbutamo_prev.csv", row.names = TRUE)
vif_allbutamo_prev
```
and leave gear out

remove amo_prev + gear
```{r vif all but amoprev + gear}
vif_allbutamo_prevgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbutamo_prevgear, "../output/bio/vif/vif_allbutamo_prevgear.csv", row.names = TRUE)
vif_allbutamo_prevgear
```

remove chl_depth
```{r vif all but chldepth}
vif_allbutchl_depth <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutchl_depth, "../output/bio/vif/vif_allbutchl_depth.csv", row.names = TRUE)

vif_allbutchl_depth <- read.csv("../output/bio/vif/vif_allbutchl_depth.csv", header = TRUE)
vif_allbutchl_depth$GVIF2Dfsq <- vif_allbutchl_depth$GVIF..1..2.Df..^2
write.csv(vif_allbutchl_depth, "../output/bio/vif/vif_allbutchl_depth.csv", row.names = TRUE)
vif_allbutchl_depth
```

remove chl_depth and gear
```{r vif all but chldepth+gear}
vif_allbutchl_depthgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutchl_depthgear, "../output/bio/vif/vif_allbutchl_depthgear.csv", row.names = TRUE)
vif_allbutchl_depthgear
```

remove ssh_surface
```{r vif all but ssh}
vif_allbutssh_surface <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutssh_surface, "../output/bio/vif/vif_allbutssh_surface.csv", row.names = TRUE)

vif_allbutssh_surface <- read.csv("../output/bio/vif/vif_allbutssh_surface.csv", header = TRUE)
vif_allbutssh_surface$GVIF2Dfsq <- vif_allbutssh_surface$GVIF..1..2.Df..^2
write.csv(vif_allbutssh_surface, "../output/bio/vif/vif_allbutssh_surface.csv", row.names = TRUE)
vif_allbutssh_surface
```

remove ssh_surface & gear
```{r vif all but ssh+gear}
vif_allbutssh_surfacegear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutssh_surfacegear, "../output/bio/vif/vif_allbutssh_surfacegear.csv", row.names = TRUE)
vif_allbutssh_surfacegear
```

remove o2_surface
```{r vif all but o2surface}
vif_allbuto2_surface <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbuto2_surface, "../output/bio/vif/vif_allbuto2_surface.csv", row.names = TRUE)

vif_allbuto2_surface <- read.csv("../output/bio/vif/vif_allbuto2_surface.csv", header = TRUE)
vif_allbuto2_surface$GVIF2Dfsq <- vif_allbuto2_surface$GVIF..1..2.Df..^2
write.csv(vif_allbuto2_surface, "../output/bio/vif/vif_allbuto2_surface.csv", row.names = TRUE)
vif_allbuto2_surface
```

remove o2_surface & gear
```{r vif all but o2surface+gear}
vif_allbuto2_surfacegear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbuto2_surfacegear, "../output/bio/vif/vif_allbuto2_surfacegear.csv", row.names = TRUE)
vif_allbuto2_surfacegear
```

@@

remove salinity_surface only
```{r vif all but salsurface}
vif_allbutsalinitysurface <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutsalinitysurface, "../output/bio/vif/vif_allbutsalinitysurface.csv", row.names = TRUE)

vif_allbutsalinitysurface <- read.csv("../output/bio/vif/vif_allbutsalinitysurface.csv", header = TRUE)
vif_allbutsalinitysurface$GVIF2Dfsq <- vif_allbutsalinitysurface$GVIF..1..2.Df..^2
write.csv(vif_allbutsalinitysurface, "../output/bio/vif/vif_allbutsalinitysurface.csv", row.names = TRUE)
vif_allbutsalinitysurface
```
and without gear

```{r vif all but salsurface+gear}
vif_allbutsalinitysurfacegear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutsalinitysurfacegear, "../output/bio/vif/vif_allbutsalinitysurfacegear.csv", row.names = TRUE)
vif_allbutsalinitysurfacegear
```

remove temp_surface only
```{r vif all but tempsurface}
vif_allbuttempsurface <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbuttempsurface, "../output/bio/vif/vif_allbuttempsurface.csv", row.names = TRUE)

vif_allbuttempsurface <- read.csv("../output/bio/vif/vif_allbuttempsurface.csv", header = TRUE)
vif_allbuttempsurface$GVIF2Dfsq <- vif_allbuttempsurface$GVIF..1..2.Df..^2
write.csv(vif_allbuttempsurface, "../output/bio/vif/vif_allbuttempsurface.csv", row.names = TRUE)
vif_allbuttempsurface

vif_allbuttempsurfacegear <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbuttempsurfacegear, "../output/bio/vif/vif_allbuttempsurfacegear.csv", row.names = TRUE)
vif_allbuttempsurfacegear

now remove temp_surface plus highly correlated
```{r vif allbut high corr + temp surface/gear/nogear}
#with gear
vif_allbuthighcorrtempsurface <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbuthighcorrtempsurface, "../output/bio/vif/vif_allbuthighcorrtempsurface.csv", row.names = TRUE)

vif_allbuthighcorrtempsurface <- read.csv("../output/bio/vif/vif_allbuthighcorrtempsurface.csv", header = TRUE)
vif_allbuthighcorrtempsurface$GVIF2Dfsq <- vif_allbuthighcorrtempsurface$GVIF..1..2.Df..^2
write.csv(vif_allbuthighcorrtempsurface, "../output/bio/vif/vif_allbuthighcorrtempsurface.csv", row.names = TRUE)
vif_allbuthighcorrtempsurface

#without gear
vif_allbuthighcorrtempsurfacegear <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbuthighcorrtempsurfacegear, "../output/bio/vif/vif_allbuthighcorrtempsurfacegear.csv", row.names = TRUE)
vif_allbuthighcorrtempsurfacegear
```

now remove temp_surface plus salinity_surface
```{r vif allbut temp + salinity surface/gear/nogear}
#with gear
vif_allbuttempsalsurface <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbuttempsalsurface, "../output/bio/vif/vif_allbuttempsalsurface.csv", row.names = TRUE)

vif_allbuttempsalsurface <- read.csv("../output/bio/vif/vif_allbuttempsalsurface.csv", header = TRUE)
vif_allbuttempsalsurface$GVIF2Dfsq <- vif_allbuttempsalsurface$GVIF..1..2.Df..^2
write.csv(vif_allbuttempsalsurface, "../output/bio/vif/vif_allbuttempsalsurface.csv", row.names = TRUE)
vif_allbuttempsalsurface

#without gear
vif_allbuttempsalsurfacegear <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbuttempsalsurfacegear, "../output/bio/vif/vif_allbuttempsalsurfacegear.csv", row.names = TRUE)
vif_allbuttempsalsurfacegear
```

now remove temp_surface plus salinity_surface + highly correlated
```{r vif allbut high corr +  temp + salinity surface/gear/nogear}
#with gear
vif_allbuthighcorrtempsalsurface <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbuthighcorrtempsalsurface, "../output/bio/vif/vif_allbuthighcorrtempsalsurface.csv", row.names = TRUE)

vif_allbuthighcorrtempsalsurface <- read.csv("../output/bio/vif/vif_allbuthighcorrtempsalsurface.csv", header = TRUE)
vif_allbuthighcorrtempsalsurface$GVIF2Dfsq <- vif_allbuthighcorrtempsalsurface$GVIF..1..2.Df..^2
write.csv(vif_allbuthighcorrtempsalsurface, "../output/bio/vif/vif_allbuthighcorrtempsalsurface.csv", row.names = TRUE)
vif_allbuthighcorrtempsalsurface

#without gear
vif_allbuthighcorrtempsalsurfacegear <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbuthighcorrtempsalsurfacegear, "../output/bio/vif/vif_allbuthighcorrtempsalsurfacegear.csv", row.names = TRUE)
vif_allbuthighcorrtempsalsurfacegear
```

@@

```{r}
vif_allbutnaoamo <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear, data = presab))
write.csv(vif_allbutnaoamo, "../output/bio/vif/vif_allbutnaoamo.csv", row.names = TRUE)

vif_allbutnaoamo <- read.csv("../output/bio/vif/vif_allbutnaoamo.csv", header = TRUE)
vif_allbutnaoamo$GVIF2Dfsq <- vif_allbutnaoamo$GVIF..1..2.Df..^2
write.csv(vif_allbutnaoamo, "../output/bio/vif/vif_allbutnaoamo.csv", row.names = TRUE)
vif_allbutnaoamo
```
not much diff..
